Final Project Code ALL

### 2022-12-08 Draft_Final_Code.Rmd
Almost finalized linear regression code

## Linear Regression Model

@@

### Fit Preliminary Model

@@

### Obtain Evaluation Metrics for Preliminary Model

@@

### Perform LASSO for Subset Selection

@@

best_penalty <- select_best(tune_res, metric = 'rmse') # choose penalty value

@@

### Fit Final Model

@@

### Visualize Residuals

@@

## GAM Using Splines

@@

gam_mod <- fit(gam_spec,
    Gross ~ s(Runtime) + IMDB_Rating + s(Meta_score) + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

@@

@@

gam_mod %>% pluck('fit') %>% mgcv::gam.check()

gam_mod %>% pluck('fit') %>% summary() 
```

@@

gam_mod %>% pluck('fit') %>% plot()
```

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6)

@@

```{r}
# Use the edf (round to integer) from above to create a recipe by adding step_ns() for the variables you want model with a non-linear relationship and do CV. Compare this model to one without any splines.

@@

cv_output <- fit_resamples( 
  spline_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

cv_output %>% collect_metrics()

@@

@@

# Variable Importance for GAM

```{r}
# Lasso Model Spec with tune
gam_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

@@

@@

# Fit Final Model

@@

# Fit Final Model
final_wf1 <- finalize_workflow(lasso_wf_tune1, best_penalty) # incorporates penalty value to workflow

final_fit1 <- fit(final_wf1, data = imdb_clean)

tidy(final_fit1)
```

final_fit1 %>% tidy() %>% filter(estimate != 0)
```

# Visualize Residuals

# Visualize Residuals
lasso_mod_out1 <- new_mod %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out1, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

# Comparing Model Performance

@@

```{r}
new_modcv %>% collect_metrics() #GAM
```

@@

@@

# Compute a distance matrix on the scaled data
dist_mat_scaled <- dist(scale(imdb_hc_sub))
```

```{r}
imdb_hc_avg <- hclust(dist_mat_scaled, method = "average")

@@

```{r}
imdb_clean <- imdb_clean %>%
    mutate(
        hclust_height = factor(cutree(imdb_hc_avg, h = 1.25)), # Cut at height (h) 3
        hclust_num = factor(cutree(imdb_hc_avg, k = 5)) # Cut into 6 clusters (k)
    )

plot(imdb_hc_avg, labels = imdb_clean$hclust_height)
```

>>>>

## Ordinary Linear Regression Model

@@

### Fit Full Model

@@

### Obtain Evaluation Metrics for Full Model

@@

## Perform LASSO for Subset Selection

@@

# Choose penalty value
best_penalty <- select_best(tune_res, metric = 'rmse')

@@

### Fit Final LASSO Model

@@

### Obtain Evaluation Metrics for Full Model

```{r}
lasso_fit %>%
  collect_metrics()
```

### Visualize Residuals for LASSO Model

@@

## GAM with Splines (TidyModels)

### Build the GAM

@@

gam_mod1 <- fit(gam_spec,
    Gross ~ s(Runtime) + s(IMDB_Rating) + Meta_score + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

@@

### Run Diagnostics

@@

gam_mod1 %>% pluck('fit') %>% mgcv::gam.check()
```

```{r}
# Diagnostics: Check to see if the number of knots is large enough

gam_mod1 %>% pluck('fit') %>% summary() 
```

### Visualization for Non-Linear Functions

@@

gam_mod1 %>% pluck('fit') %>% plot()
```

### Evaluation Metrics

```{r}
gam1_output <- gam_mod1%>% 
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

gam1_output %>%
    rmse(truth = Gross, estimate = .pred)

gam1_output %>%
    rsq(truth = Gross, estimate = .pred)
```

## GAM with Splines (Recipe)

### Build the GAM

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6) %>%
    step_ns(IMDB_Rating, deg_free = 2)

@@

@@

cv_output_spline2 <- fit_resamples( 
  spline_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

@@

### Evaluation Metrics

```{r}
cv_output_spline2 %>% collect_metrics()

### idk what this is for

@@

## GAM with Splines and LASSO

### Build and Tune Model

```{r}
# Lasso Model Spec with tune
gam_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% 
  set_mode('regression')

@@

### Collect CV Metrics and Select Best Model

@@

### Fit Final GAM LASSO Model

final_wf_gam_lasso <- finalize_workflow(lasso_wf_tune1, best_penalty) # incorporates penalty value to workflow

final_fit_gam_lasso <- fit(final_wf_gam_lasso, data = imdb_clean)

tidy(final_fit_gam_lasso)
```

final_fit_gam_lasso %>% tidy() %>% filter(estimate != 0)
```

### Evaluation Metrics

```{r}
gam_lasso_output <- fit_resamples( 
  final_wf_gam_lasso, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

gam_lasso_output %>% collect_metrics()
```

## Compare GAMs

```{r}
gam1_output %>%
    rmse(truth = Gross, estimate = .pred)

cv_output_spline2 %>% collect_metrics()

gam_lasso_output %>% collect_metrics()
```

The GAM created using TidyModels performs better than the recipe and LASSO GAMs.

### Visualize Residuals for Final GAM

@@

# Visualize Residuals
gam1_output <- new_mod %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(gam1_output, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(gam1_output, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(gam1_output, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(gam1_output, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

There does not appear to be any significant bias after analyzing the residuals.

## Compare All Linear Regression Model Performance

@@

```{r}
gam1_output %>%
    rmse(truth = Gross, estimate = .pred) #GAM
```

The GAM with Splines obtained using TidyModels performs the best with the lowest RMSE value (1.3 compared to 1.4 for the others).

@@

imdb_hc_full <- imdb_clean %>%
  select(Gross, Budget)

@@

# Compute a distance matrix on the scaled data
dist_mat_scaled <- dist(scale(imdb_hc_sub))

dist_mat_full <- dist(scale(imdb_hc_full))
```

```{r}
imdb_hc_avg <- hclust(dist_mat_scaled, method = "average")
imdb_full_avg <- hclust(dist_mat_full, method = "average")

@@

```{r}
imdb_clean <- imdb_clean %>%
    mutate(
        hclust_num = factor(cutree(imdb_full_avg, k = 3)) # Cut into 6 clusters (k)
    )

```{r}
ggplot(imdb_clean, aes(x = hclust_num, fill = Genre_1)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()
```
